Dynamics of isoflurane-induced vasodilation and blood flow of cerebral vasculature revealed by multi-exposure speckle imaging

### 2021-11-04 article.tex
Minor edits

Mice ($n = 4$, C57BL/6J, male, 4-6 months, Charles River Laboratories, Inc.) were anesthetized with medical \ce{O2} vaporized isoflurane (3\% for induction, 1-2\% for maintenance, 0.5 L/min) via nose-cone inhalation. Body temperature was maintained at 37 $^\circ$C with a feedback heating pad (DC Temperature Controller, Future Health Concepts, Inc.). Arterial oxygen saturation, heart rate, and breathing rate were monitored via pulse oximetry (MouseSTAT, Kent Scientific Corp.). Mice were placed supine in a stereotaxic frame (Kent Scientific Corp.) and subcutaneously administered carprofen (5 mg/kg) and dexamethasone (2 mg/kg) to reduce inflammation during the craniotomy procedure. The scalp was shaved and resected to expose the skull between the bregma and lambda cranial coordinates. A circular region (4 mm diameter) of the skull including the dura mater over the somatosensory cortex was removed with a dental drill (Ideal Microdrill, 0.8 mm burr, Fine Science Tools, Inc.) while under regular artificial cerebrospinal fluid (buffered pH 7.4) perfusion. A 4 mm round \#1.5 cover glass (World Precision Instruments, Inc.) was placed over the exposed brain with artificial cerebrospinal fluid filling the void. Gentle pressure was applied to the cover glass while Kwik-Sil adhesive (World Precision Instruments, Inc.) was deposited to bond the glass to the skull. A layer of Vetbond tissue adhesive (3M) was applied over the Kwik-Sil to create a sterile, air-tight seal around the craniotomy and to allow for the restoration of intracranial pressure. C\&B Metabond (Parkell, Inc.) was then used to cement a custom titanium head-plate centered around the cranial window for head-constrained awake measurements. The subjects were allowed to recover from surgery and monitored for cranial window integrity and normal behavior for at least four weeks prior to imaging. They were then exposed to head fixation during 20-30 minute sessions over 3-5 days until habituated to locomotion on a custom, low-profile linear treadmill system \cite{He:2020}.

@@

Cranial window implanted mice were head constrained on the awake imaging system and allowed to acclimate for 5-10 minutes. Periods of extended locomotion on the treadmill were manually noted throughout the experiment. The awake hemodynamic baseline was defined by acquiring 50 MESI frames, averaging the speckle contrast images across each exposure time, and computing the corresponding MESI $ICT$ image using Eq.~(\ref{eq:mesi}). A custom 3D-printed inhalation nose-cone connected to the anesthesia vaporizer and scavenger was then placed several millimeters away from the subject and used to administer medical air (0.5 L/min) with no isoflurane (0\%). Continuous MESI was initiated ($t = 0$) and used to monitor the awake subject for 10 minutes before increasing isoflurane to 2.0\% to induce anesthesia. After one minute of exposure ($t = 11$ minutes), the nose-cone was fully positioned over the subject to ensure targeted delivery of isoflurane. After two minutes of exposure ($t = 12$ minutes), a feedback heating pad (55-7030, Harvard Apparatus, Inc.) was placed beneath the subject to maintain a 37 $^\circ$C body temperature. A pulse oximeter (MouseSTAT, Kent Scientific Corp.) was also attached and used to monitor vitals and regulate the depth of anesthesia such that the heart rate was matched to within 10\% between imaging sessions for each subject \cite{Kazmi:2013hp}. The now anesthetized subject was continuously monitored with MESI until $t = 30$ minutes, at which point the anesthetized hemodynamic state was defined by acquiring an additional 50 MESI frames. The speckle contrast images were then averaged across each exposure time and used to compute the MESI $ICT$ image using Eq.~(\ref{eq:mesi}). The subject was then removed from the imaging system and allowed to recover from anesthesia on a heating pad before being transferred back into its cage. Each subject underwent three imaging sessions with at least one full day of recovery between each session.

@@

All data processing was performed using MATLAB (R2021a, MathWorks, Inc.). While head fixation can greatly minimize locomotion-associated brain movement \cite{Dombeck:2007gr}, it does not completely eliminate motion artifacts. The induction of general anesthesia also causes major changes in posture that can result in lateral displacements of the brain relative to the skull. In order to account for these shifts, the open source image registration toolbox \texttt{Elastix} \cite{Klein:2010gr} was used to rigidly align all speckle contrast frames to the baseline awake imagery prior to fitting Eq.~(\ref{eq:mesi}). The resulting aligned $ICT$ frames were then smoothed temporally with a central moving average filter ($k=5$) and rendered to video at 20 frames-per-second (8$\times$ speed). $ICT$ timecourses spanning the entire imaging session were then extracted from the aligned data from both vascular and parenchymal regions of interest (ROIs).

\subsubsection{Calculating Vascular Flux}
In order to calculate the vascular flux \cite{Kazmi:2015du}, the diameter of vessels ($d_{vessel}$) were estimated using the 5 ms exposure speckle contrast frames. Cross-sectional profiles were extracted across each vessel of interest at every timepoint and fitted to a Gaussian function in order to calculate the full-width at half-maximum (FWHM) of the distribution over time. Because LSCI is most sensitive to moving scatterers, i.e. erythrocytes, this method of estimating $d_{vessel}$ is a measure of the inner diameter of the vessel. The vascular flux was then computed by scaling the $ICT$ by $d_{vessel}$ at each timepoint. The percent change in vessel diameter and vascular flux were calculated using the average of the entire awake state for the baseline.

@@

\subsubsection{Statistical Analysis}
All statistical analyses were also performed using MATLAB. Paired sample t-tests were used to compare awake and anesthetized measurements across vessel diameter and CBF metrics. The differences between arterial, venous, and parenchymal measurements were evaluated with one-way ANOVAs using the Bonferroni method for multiple comparisons. The significance level ($\alpha$) was set to 0.05 for all statistical tests.

@@

% Table 1 - Aggregated Awake vs. Anesthetized Summary
\begin{table}[]
    \caption {
        Paired samples t-tests comparing differences between awake and anesthetized measurements across vessel diameter and blood flow metrics ($\alpha = 0.05$).
    }
    \label{tab:awakeanes}
    \centering
    \begin{tabular}{llccc} 
        \addlinespace
        \toprule
        \textbf{Metric} & \textbf{Region}  & \textbf{$\boldsymbol{t}$} & \textbf{$\boldsymbol{df}$} & \textbf{$\boldsymbol{p}$} \\
        \midrule
        Vessel Dia.     & Artery ($\mu$m)  & -6.15                     & 32                         & $\num{7.1e-7}$            \\
                        & Vein ($\mu$m)    & -3.67                     & 28                         & 0.001                     \\
        \midrule
        Blood Flow      & Artery (Flux)    & -10.75                    & 32                         & $\num{3.7e-12}$           \\
                        & Vein (Flux)      & -4.96                     & 28                         & $\num{3.1e-5}$            \\
                        & Parenchyma (ICT) & -18.44                    & 45                         & $\num{1.3e-22}$           \\
        \bottomrule
    \end{tabular}
\end{table}

@@

The systemic effects of isoflurane undermine the reliability of neurophysiological and behavioral studies because anesthesia is a fundamentally unnatural physiological state \cite{Gao:2017tw}. Imaging blood flow on the cortical surface with MESI during the induction of anesthesia directly visualizes and quantifies these large hemodynamic changes. On average, the vasodilation caused by isoflurane increased vessel diameter by 14.1\% across all trials, with arteries increasing by 15.0\% and veins increasing by 13.0\% (Fig.~\ref{fig:aggregated}a). Direct microscope measurements in fentanyl- and nitrous oxide-anesthetized rats documented a 17\% increase in arteriolar diameter and 6\% increase in venule diameter after the topical application of isoflurane \cite{Koenig:1994rn}. Similar measurements in pentobarbital-anesthetized dogs observed 10-28\% increases in the diameter of small arterioles following the inhalation of isoflurane \cite{Iida:1998th}. Both studies found that isoflurane dilated arterioles in a concentration-dependent manner. More recent measurements in mice using optical coherence tomography (OCT) angiography found that isoflurane increased artery diameter by 12-55\% and vein diameter by 14-22\%, depending on vessel branch order \cite{Rakymzhan.2021}.

@@

While LSCI has been utilized extensively with awake imaging \cite{Takuwa:2011jr,Seto:2014ga,Lu:2017bo,Balbi:2017cj,Sunil:2020ac}, this is one of the first uses of MESI in an awake animal model \cite{He:2020}. Traditional single-exposure LSCI would likely have underestimated the magnitude of the flow changes caused by isoflurane because individual exposure times are only sensitive to a narrow range of flow rates \cite{Parthasarathy:2008el}. Failing to account for the increase in multiple dynamic scattering events caused by vasodilation would also have resulted in an underestimation of the flow changes in the surface vasculature. Using the raw $ICT$ value instead of the diameter-scaled vascular flux metric in Fig.~\ref{fig:timecourse}c would have measured only a 141\% increase in the large vessel instead of 270\%. The repeatability of the day-to-day measurements shown in Fig.~\ref{fig:timecourse}d-f further demonstrate the robustness of MESI as a chronic blood flow imaging technique \cite{Kazmi:2013hp}, even when observing a fully-conscious, moving animal.

>>>>

Mice ($n = 4$, C57BL/6J, male, 4-6 months, Charles River Laboratories, Inc.) were anesthetized with medical \ce{O2} vaporized isoflurane (3\% for induction, 1-2\% for maintenance, 0.5 L/min) via nose-cone inhalation. Body temperature was maintained at 37 $^\circ$C with a feedback heating pad (DC Temperature Controller, Future Health Concepts, Inc.). Arterial oxygen saturation, heart rate, and breathing rate were monitored via pulse oximetry (MouseSTAT, Kent Scientific Corp.). Mice were placed supine in a stereotaxic frame (Kent Scientific Corp.) and subcutaneously administered carprofen (5 mg/kg) and dexamethasone (2 mg/kg) to reduce inflammation during the craniotomy procedure. The scalp was shaved and resected to expose the skull between the bregma and lambda cranial coordinates. A circular region (4 mm diameter) of the skull including the dura mater over the somatosensory cortex was removed with a dental drill (Ideal Microdrill, 0.8 mm burr, Fine Science Tools, Inc.) while under regular artificial cerebrospinal fluid (buffered pH 7.4) perfusion. A 4 mm round cover glass (\#1.5, World Precision Instruments, Inc.) was placed over the exposed brain with artificial cerebrospinal fluid filling the void. Gentle pressure was applied to the cover glass while Kwik-Sil adhesive (World Precision Instruments, Inc.) was deposited to bond the glass to the skull. A layer of Vetbond tissue adhesive (3M) was applied over the Kwik-Sil to create a sterile, air-tight seal around the craniotomy and to allow for the restoration of intracranial pressure. C\&B Metabond (Parkell, Inc.) was then used to cement a custom titanium head-plate centered around the cranial window for head-constrained awake measurements. The subjects were allowed to recover from surgery and monitored for cranial window integrity and normal behavior for at least four weeks prior to imaging. They were then exposed to head fixation during 20-30 minute sessions over 3-5 days until habituated to locomotion on a custom low-profile linear treadmill system \cite{He:2020}.

@@

Cranial window implanted mice were head constrained on the awake imaging system and allowed to acclimate for 5-10 minutes. Periods of extended locomotion on the treadmill were manually noted throughout the experiment. The awake hemodynamic baseline was defined by acquiring 50 MESI frames, averaging the speckle contrast images across each exposure time, and computing the corresponding MESI $ICT$ image using Eq.~(\ref{eq:mesi}). A custom 3D-printed inhalation nose-cone connected to the anesthesia vaporizer and scavenger was then placed several millimeters away from the subject and used to administer medical air (0.5 L/min) with no isoflurane (0\%). Continuous MESI was initiated ($t = 0$) and used to monitor the awake subject for 10 minutes before increasing isoflurane to 2.0\% to induce anesthesia. After one minute of exposure ($t = 11$ minutes), the nose-cone was fully positioned over the subject to ensure targeted delivery of isoflurane. After two minutes of exposure ($t = 12$ minutes), a feedback heating pad (55-7030, Harvard Apparatus, Inc.) was placed beneath the subject to maintain a 37 $^\circ$C body temperature. A pulse oximeter (MouseSTAT, Kent Scientific Corp.) was also attached and used to monitor vitals and regulate the depth of anesthesia such that the heart rate was matched to within 10\% between imaging sessions for each subject \cite{Kazmi:2013hp}. The now anesthetized subject was continuously monitored with MESI until $t = 30$ minutes, at which point the anesthetized hemodynamic state was defined by acquiring an additional 50 MESI frames. The speckle contrast images were then averaged across each exposure time and used to compute the anesthetized MESI $ICT$ image using Eq.~(\ref{eq:mesi}). The subject was then removed from the imaging system and allowed to recover from anesthesia on a heating pad before being transferred back into its cage. Each subject underwent three imaging sessions with at least one full day of recovery between each session.

@@

All data processing was performed using MATLAB (R2021a, MathWorks, Inc.). While head fixation can greatly minimize locomotion-associated brain movement \cite{Dombeck:2007gr}, it does not completely eliminate all motion. The induction of general anesthesia also causes major changes in posture that can result in lateral displacements of the brain relative to the skull. In order to account for these shifts, the open source image registration toolbox \texttt{Elastix} \cite{Klein:2010gr} was used to rigidly align all speckle contrast frames to the baseline awake imagery prior to fitting Eq.~(\ref{eq:mesi}). The resulting aligned $ICT$ frames were then smoothed temporally with a central moving average filter ($k=5$) and rendered to video at 20 frames-per-second (8$\times$ speed). $ICT$ timecourses spanning the entire imaging session were then extracted from the aligned data from both vascular and parenchymal regions of interest (ROIs).

\subsubsection{Calculating Vascular Flux}
In order to calculate the vascular flux \cite{Kazmi:2015du}, the diameter of vessels ($d_{vessel}$) were estimated using the 5 ms exposure speckle contrast frames. Cross-sectional profiles were extracted across each vessel of interest at every timepoint and fitted to a Gaussian function in order to calculate the full-width at half-maximum (FWHM) of the distribution over time. Because LSCI is most sensitive to moving scatterers, i.e. erythrocytes, this method of estimating $d_{vessel}$ is a measure of the inner diameter of the vessel. The vascular flux was then computed by scaling the $ICT$ by $d_{vessel}$ at each timepoint. The percent change in vessel diameter and vascular flux were calculated using the average of the entire awake timecourse for the baseline.

@@

\subsubsection{Statistical Analysis}
All statistical analyses were performed using MATLAB. Paired sample t-tests were used to compare awake and anesthetized measurements across vessel diameter and CBF metrics. The differences between arterial, venous, and parenchymal measurements were evaluated with one-way ANOVAs using the Bonferroni method for multiple comparisons. The significance level ($\alpha$) was set to 0.05 for all statistical tests.

@@

% Table 1 - Aggregated Awake vs. Anesthetized Summary
\begin{table}[]
    \caption {
        Paired samples t-tests comparing differences between awake and anesthetized measurements of aggregated vessel diameter and blood flow metrics ($\alpha = 0.05$).
    }
    \label{tab:awakeanes}
    \centering
    \begin{tabular}{llccc} 
        \addlinespace
        \toprule
        \textbf{Metric} & \textbf{Region}  & \textbf{$\boldsymbol{t}$} & \textbf{$\boldsymbol{df}$} & \textbf{$\boldsymbol{p}$} \\
        \midrule
        Vessel Dia.     & Artery ($\mu$m)  & -6.15                     & 32                         & $\num{7.1e-7}$            \\
                        & Vein ($\mu$m)    & -3.67                     & 28                         & 0.001                     \\
        \midrule
        Blood Flow      & Artery (Flux)    & -10.75                    & 32                         & $\num{3.7e-12}$           \\
                        & Vein (Flux)      & -4.96                     & 28                         & $\num{3.1e-5}$            \\
                        & Parenchyma (ICT) & -18.44                    & 45                         & $\num{1.3e-22}$           \\
        \bottomrule
    \end{tabular}
\end{table}

@@

The systemic effects of isoflurane undermine the reliability of neurophysiological and behavioral studies because anesthesia is a fundamentally unnatural physiological state \cite{Gao:2017tw}. Imaging blood flow on the cortical surface with MESI during the induction of anesthesia directly visualizes and quantifies these large acute hemodynamic changes. On average, the vasodilation caused by isoflurane increased vessel diameter by 14.1\% across all trials, with arteries increasing by 15.0\% and veins increasing by 13.0\% (Fig.~\ref{fig:aggregated}a). Direct microscope measurements in fentanyl- and nitrous oxide-anesthetized rats documented a 17\% increase in arteriolar diameter and 6\% increase in venule diameter after the topical application of isoflurane \cite{Koenig:1994rn}. Similar measurements in pentobarbital-anesthetized dogs observed 10-28\% increases in the diameter of small arterioles following the inhalation of isoflurane \cite{Iida:1998th}. Both studies found that isoflurane dilated arterioles in a concentration-dependent manner. More recent measurements in mice using optical coherence tomography (OCT) angiography found that isoflurane increased artery diameter by 12-55\% and vein diameter by 14-22\%, depending on vessel branch order \cite{Rakymzhan.2021}.

@@

While LSCI has been utilized extensively with awake imaging \cite{Takuwa:2011jr,Seto:2014ga,Lu:2017bo,Balbi:2017cj,Sunil:2020ac}, this is one of the first uses of MESI in an awake animal model \cite{He:2020}. Traditional single-exposure LSCI would likely have underestimated the magnitude of the flow changes caused by isoflurane because individual exposure times are only sensitive to a narrow range of flow rates \cite{Parthasarathy:2008el}. Failing to account for the increase in multiple dynamic scattering events caused by vasodilation would also have resulted in an underestimation of the flow changes in the surface vasculature. For example, using the raw $ICT$ value instead of the diameter-scaled vascular flux metric in Fig.~\ref{fig:timecourse}c would have measured only a 141\% increase in the large vessel instead of 270\%. The repeatability of the day-to-day measurements shown in Fig.~\ref{fig:timecourse}d-f further demonstrate the robustness of MESI as a chronic blood flow imaging technique \cite{Kazmi:2013hp}, even when observing a fully-conscious, moving animal.
